Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-12-05 ms/spaceness.tex
actual final edits

Following recent studies that showed strong signals for dispersal and demography in the distribution of shared haplotype block lengths \citep[e.g.,][]{Ringbauer2017,Baharian2016}, we also calculated various summaries of the distribution of pairwise identical-by-state (IBS) block lengths among sampled chromosomes,
defined to be the set of distances between adjacent sites that differ between the two chromosomes. \revpoint{1}{1}
The full distribution of lengths of IBS tracts for each pair of chromosomes was first calculated with a custom python function. 
%\url{https://github.com/cjbattey/spaceness/blob/master/scripts/slimtools.py}. 
We then calculated the first three moments of this distribution (mean, variance, and skew) and the number of blocks over $10^6$ base pairs both for each pair of individuals and for the full distribution across all pairwise comparisons. 
We then calculated correlation coefficients between spatial distance and each moment of the pairwise IBS tract distribution. Because more closely related individuals on average share longer haplotype blocks we expect that spatial distance will be negatively correlated with mean haplotype block length, and that this correlation will be strongest (i.e., most negative) when dispersal is low. The variance, skew, and count of long haplotype block statistics are meant to reflect the relative length of the right (upper) tail of the distribution, which represents the frequency of long haplotype blocks, and so should reflect recent demographic events \citep{chapman2002effect}. For a subset of simulations, we also calculated cumulative distributions for IBS tract lengths across pairs of distant (more than 48 map units) and nearby (less than 2 map units) individuals. Last, we examined the relationship between allele frequency and the spatial dispersion of an allele by calculating the average distance between individuals carrying each derived allele.

>>>>

Following recent studies that showed strong signals for dispersal and demography in the distribution of shared haplotype block lengths \citep[e.g.,][]{Ringbauer2017,Baharian2016}, we also calculated various summaries of the distribution of pairwise identical-by-state (IBS) block lengths among sampled chromosomes,
defined to be the set of distances between adjacent sites that differ between the two chromosomes. \revpoint{1}{1}
The full distribution of lengths of IBS tracts for each pair of chromosomes was first calculated with a custom python function. 
%\url{https://github.com/cjbattey/spaceness/blob/master/scripts/slimtools.py}. 
We then calculated the first three moments of this distribution (mean, variance, and skew) and the number of blocks over $10^6$ base pairs both for each pair of individuals and for the full distribution across all pairwise comparisons. 
We then calculated correlation coefficients between spatial distance and each moment of the pairwise IBS tract distribution. Because more closely related individuals on average share longer haplotype blocks we expect that spatial distance will be negatively correlated with mean haplotype block length, and that this correlation will be strongest (i.e., most negative) when dispersal is low. The variance, skew, and count of long haplotype block statistics are meant to reflect the relative length of the right (upper) tail of the distribution, which represents the frequency of long haplotype blocks, and so should reflect recent demographic events \citep{chapman2002effect}. For a subset of simulations, we also calculated cumulative distributions for IBS tract lengths across pairs of distant (more than 40 map units) and nearby (less than 10 map units) individuals. Last, we examined the relationship between allele frequency and the spatial dispersion of an allele by calculating the average distance between individuals carrying each derived allele.
